Pre-Trained Denoising Autoencoders Long Short-Term Memory Networks as probabilistic Models for Estimation of Distribution Genetic Programming

### 2023-01-09 paper.Rmd
Mon Jan  9 16:00:17 CET 2023: update-remote.sh

The pre-training strategy implemented for all experiments in this thesis is visualized as another flowchart in figure §.
The main difference to regular DAE-GP is the inclusion of an initial pre-training phase where a separate population $\hat{P}$ is first initialized and then randomly split by half into $\hat{P_{train}}$ and $\hat{P_{test}}$. 
A DAE-LSTM model $\hat{M}$ is then trained to learn the properties of $\hat{P_{train}}$ using an early stopping training mode where we stop training as soon as the validation error for $\hat{P_{test}}$ converges. 
After the training for $\hat{M}$ is stopped, the current state of $\hat{M}$ is frozen and the optimized trainable parameters $\theta_{\hat{M}}$ are saved before terminating the pre-training phase to start the DAE-GP phase. 
During the DAE-GP phase, the only difference to the traditional DAE-GP algorithm described in figure § is during model building: We initialize $M_g$ with $\theta_{\hat{M}}$ before training it to learn the properties of $X_g$.

>>>>

The pre-training strategy implemented for all experiments in this thesis is visualized as another flowchart in figure §.
The main difference to regular DAE-GP is the inclusion of an initial pre-training phase where a separate population $\hat{P}$ is first initialized and then randomly split by half into $\hat{P}_{train}$ and $\hat{P}_{test}$. 
A DAE-LSTM model $\hat{M}$ is then trained to learn the properties of $\hat{P}_{train}$ using an early stopping training mode where we stop training as soon as the validation error for $\hat{P}_{test}$ converges. 
After the training for $\hat{M}$ is stopped, the current state of $\hat{M}$ is frozen and the optimized trainable parameters $\theta_{\hat{M}}$ are saved before terminating the pre-training phase to start the DAE-GP phase. 
During the DAE-GP phase, the only difference to the traditional DAE-GP algorithm described in figure § is during model building: We initialize $M_g$ with $\theta_{\hat{M}}$ before training it to learn the properties of $X_g$.

### 2023-01-09 slides.Rmd
Mon Jan  9 15:00:17 CET 2023: update-remote.sh
Mon Jan  9 16:00:17 CET 2023: update-remote.sh

* Metaheuristik basierend auf genetischer Programmierung (GP)
* Erweiterung von GP durch künstliche, neuronalen Netzen zur Optimierung des Suchverhaltens
* Variante des Estimation of Distribution-GP (EDA-GP)

@@

## Implementation

>>>>

* Metaheuristik basierend auf genetischer Programmierung (GP)
* Erweiterung von GP durch künstliche, neuronalen Netzen zur Optimierung des Suchverhaltens
* Variante des Estimation of Distribution-GP (EDA-GP)
* Der Einsatz einer Pre-Training Strategie wurde in mehreren Publikationen als möglicher Weg für eine weitere Optimierung vorgeschlagen^[@dae-gp_2022_symreg]^[@daegp_explore_exploit)

@@

## Pre-Training für Denoising Autoencoders^[@pmlr-v5-erhan09a]

### Positive Wirkung von Pre-Training auf DAE

1. Gesteigerte Modell Performance (sinkender Testfehler)
2. Besserer Generalisierungsfähigkeit
3. Erhöhter Robustness des Algorithmus (sinkende Varianz des Testfehlers)

### Einfluss der Modell Architektur

* Positiver Effekt steigt mit zunehmender Komplexitität des DAE
* Je mehr versteckte Layer oder die Anzahl an Neuronen pro verstecktem Layer vorhanden sind, desto besserer Effekt des Pre-Trainings
* Für sehr kleine DAE, zeigt Pre-Training jedoch inverse, negative Auswirkung auf die Modell Performance

# Implementation

## Überblick

Gewählte Pre-Training Strategie:

* (Klassisches) Pre-Training: Einmaliges Training eines DAE-LSTM mit einer großen Population aus Lösungskandidaten $\hat{P}_{train}$

* Trainingsmethode Early Stopping: Abbruch des Trainings sobald der Testfehler für eine seperate Population $\hat{P}_{test}$ konvergiert

**Ausschluss anderer Pre-Training Strategien wie Re-Use Learning, Few-Shot Learning**

## Pre-Trained DAE-GP Ablauf

![Flowchart - Pre-Trained DAE-GP](./img/flowcharts/pt-dae-gp.png){height=75%}
